---
title: "hw06 - yavyx"
author: "Javier Castillo-Arnemann"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As usual, we'll start by loading the necessary packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
library(broom)
library(gapminder)
```

##Working with a nested data frame

We will first create a nested Gapminder dataframe to be able to run more complicated analysis in each group of interest (country and continent) than would only be possible with the `group_by()` function.
```{r}
nested_gapminder <- gapminder %>%
  group_by(continent, country) %>%
  nest()
```

Let's take a look at this new dataframe
```{r}
nested_gapminder
```
There is a row for every country, and a tibble corresponding to each one. Let's see the information contained in each tibble.

```{r}
nested_gapminder$data[1]
```

Here we can see the information that corresponds to the first country, Afghanistan.

Next, we'll define functions to get linear least-squares and robust regressions to be able to compare how accurately they model the data. We'll model how population changes through time in each country.

```{r}
linear_model <- function(data) {
  offset <- min(data$year)
  lm(pop ~ I(year - offset), data)
}

robust_model <- function(data) {
  offset <- min(data$year)
  MASS::rlm(pop ~ I(year - offset), data) 
}
```

The `I()` function tells R to interpret the minus sign as an arithmetic operator instead of a formula operator. The offset is to ensure that our model has the right intercept, and corresponds to the lowest year value in each dataset.

We will test our functions with the first country and to confirm it's working as it should before iterating through every country.


```{r}
linear_model(nested_gapminder$data[[1]])
robust_model(nested_gapminder$data[[1]])
```

It weems to be working, so we will add a new column to our nested dataframe with the information of each model.

```{r}
nested_gapminder <- nested_gapminder %>%
  mutate(linear_fit = map(data, linear_model),
         robust_fit = map(data, robust_model))  
```

After doing this, we can use the `broom` package to see how every data point and extract the data that we need from each model. For this, we use the `augment()` function that gives us the residuals for every data point, and the `glance()` function that summarizes how well the model fits the data.
```{r}
nested_gapminder <- nested_gapminder %>%
    mutate(linear_augment = map(linear_fit, augment),
           robust_augment = map(robust_fit, augment),
           linear_glance = map(linear_fit, glance),
           robust_glance = map(robust_fit, glance)) 

nested_gapminder$linear_augment[[1]]
nested_gapminder$robust_augment[[1]]
nested_gapminder$linear_glance[[1]]
nested_gapminder$robust_glance[[1]]
```

Now we can unnest the dataframe to perform some analysis on the residuals.

```{r}
gapminder_residuals <- nested_gapminder %>%
  dplyr::select(continent, country, linear_augment, robust_augment) %>%
  unnest(linear_augment, robust_augment)
```

```{r}
mean(gapminder_residuals$.resid)
mean(gapminder_residuals$.resid1)

max_residual_countries <- gapminder_residuals %>%
  group_by(continent) %>%
  filter(.std.resid == max(.std.resid)) #Find datapoints with max residuals
  
gapminder_residuals %>%
  group_by(continent) %>%
  filter(.resid1 == max(.resid1))
```



```{r}
gapminder %>%
  filter(country == "Dominican Republic") %>%
  ggplot(aes(pop, year-1952)) +
    geom_point() +
    geom_smooth(method = "lm")
```




