---
title: "hw06 - yavyx"
author: "Javier Castillo-Arnemann"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As usual, we'll start by loading the necessary packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
library(broom)
library(gapminder)
library(MASS)
```

##Working with a nested data frame

We will first create a nested Gapminder dataframe to be able to run more complicated analysis in each group of interest (country and continent) than would be possible with the `group_by()` function.
```{r}
nested_gapminder <- gapminder %>%
  group_by(continent, country) %>%
  nest()
```

Let's take a look at this new dataframe
```{r}
nested_gapminder
```
There is a row for every country, and a tibble corresponding to each one. Let's see the information contained in each tibble.

```{r}
nested_gapminder$data[1]
```

Here we can see the information that corresponds to the first country, Afghanistan.

Next, we'll define functions to get linear least-squares and robust regressions to be able to compare how accurately they model the data. We'll model how population life expeectancy through time in each country.

```{r}
linear_model <- function(data) {
  offset <- min(data$year)
  lm(lifeExp ~ I(year - offset), data)
}

robust_model <- function(data) {
  offset <- min(data$year)
  rlm(lifeExp ~ I(year - offset), data) 
}
```

The `I()` function tells R to interpret the minus sign as an arithmetic operator instead of a formula operator. The offset is to ensure that our model has the right intercept, and corresponds to the lowest year value in each dataset.

We will test our functions with the first country and to confirm it's working as it should before iterating through every country.


```{r}
linear_model(nested_gapminder$data[[1]])
```

It weems to be working, so we will add a new column to our nested dataframe with the information from each model.

```{r}
nested_gapminder <- nested_gapminder %>%
  mutate(linear_fit = map(data, linear_model),
         robust_fit = map(data, robust_model))  
```

After doing this, we can use the `broom` package to see how every data point is modeled and extract the data that we want from each model. For this, we use the `augment()` function that gives us the residuals for every data point.
```{r}
nested_gapminder <- nested_gapminder %>%
    mutate(linear_augment = map(linear_fit, augment),
           robust_augment = map(robust_fit, augment))

nested_gapminder$linear_augment[[1]] #checking contents of first tibble
nested_gapminder$robust_augment[[1]] #checking contents of first tibble
```

Now we can unnest the dataframe to perform some analysis on the residuals.

```{r}
gapminder_residuals <- nested_gapminder %>%
  dplyr::select(continent, country, linear_augment, robust_augment) %>%
  unnest(linear_augment, robust_augment)
```

```{r}
gapminder_residuals <- gapminder_residuals %>%
  mutate(residuals_diff = .resid  - .resid1) #Get the difference in residuals 

knitr::kable(gapminder_residuals %>%
  filter(residuals_diff == max(residuals_diff))) #Find the country with the highest residual difference between the two models
  
knitr::kable(gapminder_residuals %>%
  arrange(residuals_diff) %>%
  head()) #Arrange the data in order of residual difference
```

It is Swaziland, so let's plot its information to take a closer look at what's going on.

```{r}
gapminder %>%
  filter(country == "Swaziland") %>%
  ggplot(aes(year, lifeExp)) +
    geom_point() +
    geom_smooth(method = "lm", color = "blue") +
    geom_smooth(method = "rlm", color = "red") +
    labs(title = "Life Expectancy in Swaziland",
         x = "Year",
         y = "Life Expectancy") +
    theme_bw()
```

Neither model really gets close to the actual data, because it doesn't seem to behave in a linear way. Therefore, a quadratic model makes more sense in this case:

```{r}
gapminder %>%
  filter(country == "Swaziland") %>%
  ggplot(aes(x = year, y = lifeExp)) +
    geom_point() +
    geom_smooth(method = "lm", formula = y ~ x + I(x^2), color = "blue") + #specify qudratic model
    labs(title = "Life Expectancy in Swaziland",
         x = "Year",
         y = "Life Expectancy") +
    theme_bw()
```

##References
- [Robust Regression Using R](http://www.alastairsanderson.com/R/tutorials/robust-regression-in-R/)
- [MASS:rlm Documentation](https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/rlm.html)


